prolfqua&colon; A Comprehensive R-package for Proteomics Differential Expression Analysis

### 2022-10-21 inst/MyArticle/paper.Rmd
cod snippet cosmetics

linfct <- prolfqua::linfct_from_model(mm$modelDF$linear_model[[1]],
                                      as_list = FALSE)
mC <- prolfqua::linfct_matrix_contrasts(linfct, Contrasts)

@@

## read MQ peptide.txt and annotation table
startdata <- prolfqua::tidyMQ_Peptides(system.file(
  "samples/maxquant_txt/tiny2.zip", package = "prolfqua"))
annot <- readxl::read_xlsx(system.file(
  "samples/maxquant_txt/annotation_Ionstar2018_PXD003881.xlsx", package = "prolfqua"))
startdata <- dplyr::inner_join(annot, startdata, by = "raw.file")
## create MaxQuant configuration
config <- prolfqua::create_config_MQ_peptide()
## specify explanatory variable
config$table$factors[['dilution.']] = "sample"
## create R6 object
lfqpep <- prolfqua::LFQData$new(startdata, config, setup = TRUE) 
## remove observation with 0 intensity and filter for 2 peptides per protein
lfqpep$remove_small_intensities()$filter_proteins_by_peptide_count()
##  transform intensities
lfqpep <- lfqpep$get_Transformer()$log2()$robscale()$lfq
lfqpep$rename_response("log_peptide_abundance")
agr <- lfqpep$get_Aggregator()
lfqpro <- agr$medpolish()
lfqpro$rename_response("log_protein_abundance")
## plot panels A-D
pl <- lfqpep$get_Plotter()
panelA <- pl$intensity_distribution_density() +
  ggplot2::labs(tag = "A") + ggplot2::theme(legend.position = "none")
panelB <- agr$plot()$plots[[54]] + ggplot2::labs(tag = "B")
panelC <- lfqpro$get_Stats()$violin() + ggplot2::labs(tag = "C")
pl <- lfqpro$get_Plotter()
panelD <- pl$boxplots()$boxplot[[54]] + ggplot2::labs(tag = "D")
ggpubr::ggarrange(panelA, panelB, panelC, panelD)

@@

```{r figure2, fig.cap="(ref:figure2)", echo=TRUE, out.width = '90%', fig.width = 8, fig.height = 5}
# specify differences among groups
contrasts <- c(
  "dilution_(9/7.5)_1.2" =   "dilution.e - dilution.d",
  "dilution_(7.5/6)_1.25" =   "dilution.d - dilution.c"
)
## fit model
lmmodel <- paste(lfqpro$response()," ~ dilution.")
modelFunction <- prolfqua::strategy_lm( lmmodel, model_name = "lm")
models <- prolfqua::build_model(lfqpro, modelFunction)

## compute contrasts from linear model
contr <- prolfqua::Contrasts$new(models, contrasts, modelName = "lm")
# compute contrasts using imputation imputation
conI <- prolfqua::ContrastsSimpleImpute$new( lfqpro, contrasts, modelName = "imp")
# merge contrasts, to obtain differences for all proteins
contrasts <- prolfqua::merge_contrasts_results(prefer = contr, add = conI)
# moderate t-statistics and p-values
prolfqua::ContrastsModerated$undebug("get_contrasts")
contrasts <-  contrasts$merged |> prolfqua::ContrastsModerated$new() 
## visualize results in panels
pl <- contrasts$get_Plotter() 
histpval <- pl$histogram()$p.value + ggplot2::labs(tag = "A")               
maplot <- pl$ma_plot(rank = TRUE,legend = FALSE) + ggplot2::labs(tag = "C")
volcano <- pl$volcano()$FDR +
   ggplot2::theme(legend.position = "bottom") +
   ggplot2::labs(tag = "B")

>>>>

linfct <- prolfqua::linfct_from_model(mm$modelDF$linear_model[[1]],
                                      as_list = FALSE)
mC <- prolfqua::linfct_matrix_contrasts(linfct, Contrasts)
## see Table 3 for content of mC
```

@@

## read MQ peptide.txt and annotation table
startdata <- prolfqua::tidyMQ_Peptides(system.file(
  "samples/maxquant_txt/tiny2.zip", package = "prolfqua"))
annot <- readxl::read_xlsx(system.file(
  "samples/maxquant_txt/annotation_Ionstar2018_PXD003881.xlsx", package = "prolfqua"))
startdata <- dplyr::inner_join(annot, startdata, by = "raw.file")

## create MaxQuant configuration
config <- prolfqua::create_config_MQ_peptide()

## specify explanatory variable
config$table$factors[['dilution.']] = "sample"

## create R6 object
lfqpep <- prolfqua::LFQData$new(startdata, config, setup = TRUE)

## remove observation with 0 intensity and filter for 2 peptides per protein
lfqpep$remove_small_intensities()$filter_proteins_by_peptide_count()

## transform intensities
lfqpep <- lfqpep$get_Transformer()$log2()$robscale()$lfq
lfqpep$rename_response("log_peptide_abundance")
agr <- lfqpep$get_Aggregator()
lfqpro <- agr$medpolish()
lfqpro$rename_response("log_protein_abundance")

## plot panels A-D
pl <- lfqpep$get_Plotter()
panelA <- pl$intensity_distribution_density() +
  ggplot2::labs(tag = "A") + ggplot2::theme(legend.position = "none")
panelB <- agr$plot()$plots[[54]] + ggplot2::labs(tag = "B")
panelC <- lfqpro$get_Stats()$violin() + ggplot2::labs(tag = "C")
pl <- lfqpro$get_Plotter()
panelD <- pl$boxplots()$boxplot[[54]] + ggplot2::labs(tag = "D")
ggpubr::ggarrange(panelA, panelB, panelC, panelD)
```

@@

```{r figure2, fig.cap="(ref:figure2)", echo=TRUE, out.width = '90%', fig.width = 8, fig.height = 5}
## specify differences among groups
contrasts <- c(
  "dilution_(9/7.5)_1.2" =   "dilution.e - dilution.d",
  "dilution_(7.5/6)_1.25" =   "dilution.d - dilution.c"
)
## fit model
lmmodel <- paste(lfqpro$response()," ~ dilution.")
modelFunction <- prolfqua::strategy_lm( lmmodel, model_name = "lm")
models <- prolfqua::build_model(lfqpro, modelFunction)

## compute contrasts from linear model
contr <- prolfqua::Contrasts$new(models, contrasts, modelName = "lm")

## compute contrasts using imputation imputation
conI <- prolfqua::ContrastsSimpleImpute$new( lfqpro, contrasts, modelName = "imp")

## merge contrasts, to obtain differences for all proteins
contrasts <- prolfqua::merge_contrasts_results(prefer = contr, add = conI)

## moderate t-statistics and p-values
prolfqua::ContrastsModerated$undebug("get_contrasts")
contrasts <-  contrasts$merged |> prolfqua::ContrastsModerated$new()

## visualize results in panels
pl <- contrasts$get_Plotter() 
histpval <- pl$histogram()$p.value + ggplot2::labs(tag = "A")               
maplot <- pl$ma_plot(rank = TRUE,legend = FALSE) + ggplot2::labs(tag = "C")
volcano <- pl$volcano()$FDR +
   ggplot2::theme(legend.position = "bottom") +
   ggplot2::labs(tag = "B")
